Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble

### 2022-10-11 paper/paper.Rmd
read through for typos

---
title: Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble
preprint: true
author: 
  - name: Sam Abbott
    affiliation: 1
    corresponding: true
    email: sam.abbott@lshtm.ac.uk
  - name: Katharine Sherratt
    affiliation: 1
  - name: Nikos Bosse
    affiliation: 1
  - name: Hugo Gruson
    affiliation: 1
  - name:  Johannes Bracher
    affiliation: 2
  - name: Sebastian Funk
    affiliation: 1
affiliation:
  - code: 1
    address: London School of Hygiene and Tropical Medicine, London, UK
  - code: 2
    address: Chair of Statistics and Econometrics, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
abstract: > 
  Multi-model and multi-team ensemble forecasts have become widely used in order to generate reliable short-term predictions of infectious disease spread. Notably, various public health agencies have used them to leverage academic disease modelling during the COVID-19 pandemic. However, ensemble forecasts are difficult to interpret and require extensive effort from numerous participating groups as well as a coordination team. In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models. Here we used observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience to identify a set of properties present in current ensemble forecasts. We then developed a parsimonious forecast model with the aim of mirroring these properties. We assess forecasts generated from this model in real time over a six-month period (the 15th of January 2022 to the 19th of July 2022) and for multiple European countries. We focused on forecasts of cases one to four weeks ahead and compared them to those by the European forecast hub ensemble. We find that the surrogate model behaves qualitatively similarly to the ensemble in many instances, though with increased uncertainty and poorer performance around periods of peak incidence as measured by the Weighted Interval Score. The performance differences, however, seem to be partially driven by a subset of time points, and the proposed model appears better probabilistically calibrated than the ensemble. We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble, but more work is needed to understand the implicit epidemiological model that it represents. 
bibliography: library.bib
link-citations: true
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article
    extra_dependencies: ["float"]
    keep_tex: true
---

@@

Whilst there is robust and consistent evidence that multi-team ensemble forecasts provide reliable and performant forecasts across domains [@Reich2022-jo] they also have a range of downsides.
The most significant is the difficulty in interpreting them. This relates both to the underlying mechanisms for the forecasts they produce and to understanding if and when their behaviour is desirable. 
This impacts users' trust, how easily ensemble performance can be improved, and how easily contributor forecasts can be improved.
Forecasts from these ensembles also require considerable resource cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
Additional challenges with maintaining multi-team collaborations can include providing detailed feedback to those contributing forecasts that would allow them to improve their forecast approaches, providing incentives for forecasters to continue to contribute and adjust their models to changing conditions, and difficulty improving the quality of the ensemble by learning from past predictive performance [@Sherratt2022-ue].
Each of these issues may impact the long-term quality of the resulting forecasts and have implications for end-users. 
Little progress has so far been made in mitigating these downsides or in improving access to the high-quality and robust forecasts they seek to generate to geographies without coverage or for other infectious diseases. 
There has also been limited critical feedback on the structure of forecasting ensembling projects for infectious disease epidemiology and little evaluation of the effort required to produce them relative to their benefits for improving forecast performance.

In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysics [@Vernon2014-ov], emulation approaches have been used to circumvent resource requirement issues for complex models by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
These approaches generally take the same inputs as the models they seek to emulate and then are trained based on the output from those models.
In the context of epidemiological models, non-parametric emulation has been used to allow the rapid exploration of the parameter space of complex models that would otherwise be resource prohibitive [@Iskauskas2022-fz; @Charles2022-ds].
Replicating forecasts from resource intensive models may also be a promising avenue of future research. 
However, as an approach it may be less useful for resolving some of the issues of multi-team and multi-model forecasts as it does not provide interpretability, key for stakeholder take-up. Additionally, it is not clear how the methods commonly used perform out of sample, and emulating a quantile based forecast is an unaddressed methodological challenge as far as we are aware.

@@

In order to achieve these aims, we evaluate an initial attempt at developing a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles based on a set of clear assumptions.
We submitted this model to the European Forecast Hub and here we evaluate its performance relative to the Hub ensemble.
In this work, we first define the model and summarise its implementation, with a focus on minimal resource use and reproducibility as a GitHub Actions workflow [@GitHubActionsAbout].  
We then evaluate its real-time performance in comparison to the European Forecast Hub ensemble by visualising forecasts, using the weighted interval score [@bracherEvaluatingEpidemicForecasts2021], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where it performs less well.
Finally, we summarise our findings, discuss their implications, and highlight areas where more work is needed.
We aim for this work to highlight some of the potential implicit assumptions of current COVID-19 Forecast Hub ensembles, provide a sensible, low-resource, surrogate modelwhere large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.

@@

To understand the behaviour of the Forecast Hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team. 
Each hub targets a range of metrics, including test-positive reported cases, reported deaths, and hospitalisations; has a specific geographic focus; and asks for weekly forecasts (using epidemiological weeks i.e. Sunday to Saturday) over a time horizon of a few weeks.
Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
Here, we focus on reported cases and primarily on the European Forecast Hub but our observations hold, in our view, across COVID-19 Forecast Hubs and to a lesser degree targets.
We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator for other metrics such as hospitalisations [@Meakin2022-hw], and they are generally the most challenging to predict [@Sherratt2022-ue].
In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models, mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@Srivastava2020-qv; @Li2021-cj], semi-mechanistic approaches that blend both of these approaches [@Castro2021-lj; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-jn; @Adamik2020-if], and human insight based forecast models that may also include elements of other methods [@Karlen2020-pt; @Bosse2022-hm].
Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].

We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub [@Sherratt2022-ue; @EuroHub] from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth` and defined in the following section).
We included all locations covered by the European forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@JHU].
We used the latest available observed data as of the 1st of September 2022 (commit `f6922c3e4bdcb055abcbba8e73472afacac4cf40` from [@EuroHubGitHub]).
Incidence was aggregated by epidemiological week (i.e. Sunday through Saturday).
As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
To account for this we followed the practice of the European forecasting hub project in excluding forecasts made using anomalous truth data in the week of the forecasts production and excluding forecasts for target weeks with anomalous data [@Sherratt2022-ue].
We defined anomalous data based on the implementation used by [@EuroHub] where a datapoint is considered anomalous if a future revision alters it by more than 5%.

@@

We have contributed a range of forecasts to COVID-19 forecast collaborations, generally focused on semi-mechanistic statistical methods and human insight-based forecasts.
Our forecast submissions have not systematically over- or under-performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
Based on our observations our forecasts have generally captured the current trend relatively well but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns. 
Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European forecasting hub [@AzurePricing]. 
This resource usage is likely beyond the capacity of many interested in having access to state of the art short-term forecasts of infectious disease.
In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
We found that our human insight-based forecasts outperformed our model-based forecasts on average. 
This was particularly the case when forecasting cases and at longer forecast horizons.
We hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].

@@

The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models. Instead we make use of a negative binomial observation model allowing the inclusion of overdispersion.
This choice is motivated by our belief that the underlying transmission process is an exponential discrete one and therefore a count error model with a log link function, where variance is linked to the mean, is a sensible choice. 
We suggest that part of the reason the hub ensembles exhibit such sharpness is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021]. Our set of assumptions and simplifications are summarised in Table \@ref(tab:assumptions).

@@

In our study period, incidence rates across European nations and in the UK were primarily driven by the spread of novel subvariants of concern related to the Omicron variant and changes in population susceptibility.
Many countries, such as the UK, saw large BA.1 waves in January that then resulted in declining incidence rates through February (Figure \@ref(fig:vis-forecasts)).
From late February through to the end of May, most nations saw another wave driven by BA.2. This wave typically saw lower reported incidence rates, and was characterised by a lower peak than the BA.1 wave with a more gradual decrease in incidence.
The end of our study period was dominated by the gradual take-over of the BA.4/BA.5 subvariants that again had a lower peak and lower absolute growth rates. Unlike earlier periods in the pandemic, our study period did not see the use of new non-pharmaceutical interventions (NPIs) in response to increasing COVID-19 incidence in most locations.
In addition, ascertainment rates likely reduced over time in most locations due to reductions in routine testing, and reductions in test availability. Whilst both the reduced use of NPIs and testing generally occurred across nations our study period also marked an increase in the heterogeneity of the response to the COVID-19 pandemic with nations changing policy at different times and to different degrees.
This is in contrast to the early COVID-19 pandemic response for which most nations took similar actions at similar times.

We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European forecasting hub (nations of the European Union, the European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
In total `r nrow(metadata)` forecasts were made across all locations, with `r forecast_dates` unique forecast dates and `r models` independent forecast models (including the European hub baseline model).
Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model.
Of the remaining models submitted `r length(local_models)` were submitted in only one location.
Single location models were clustered in a few locations, particularly in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project [@Sherratt2022-ue]).
Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
`r nrow(partial_models)` models were submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.

@@

```{r vis-forecasts, fig.cap = "a.) Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom, by forecast horizon (one and four weeks) for the surrogate model (orange) and forecast ensemble (green). 30\\%, 60\\%, and 90\\% prediction intervals are shown. The black line and points are the notified cases as of the date of data extraction rather than those available at the time. b.) A replicate of a.) but with incidence rates on the log scale. c.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom on the log scale.", out.width = "98%", fig.height = 8, fig.width = 12}
p <- (
  plot_forecast_custom(forecasts, log = FALSE) |
  plot_forecast_custom(forecasts, log = TRUE) +
  labs(y = "") |
  plot_wis(wis, locs)
) +
  plot_layout(widths = c(2, 2, 1), guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre"
  )
suppressWarnings(print(p))
```

@@

```{r plot-coverage, fig.height = 8, fig.width = 8, out.width = "98%", fig.cap = "a.) Empirical coverage of the surrogate (orange) and ensemble (green) forecasts at the 90\\%, 60\\%, and 30\\% prediction intervals stratified by forecast horizon. Ideally, a well-calibrated forecast should have empirical coverage for a given prediction interval that equals the nominal level of the interval (i.e., 30\\%, 60\\% and 90\\%, respectively). b.) Empirical coverage by quantile for both the surrogate and ensemble forecasts. A well-calibrated forecast should have empirical quantiles that match the theoretical ones. The green area of this figure corresponds to conservative forecasts. c.) Median relative weighted interval score by quantile and forecast horizon for the surrogate forecast compared to the ensemble forecast. d.) Bias of the ensemble and surrogate forecasts stratified by horizon."}
p <- (
  (
    (
      plot_coverage_range(coverage, ranges) +
       guides(
        fill = guide_none(), point_color = guide_none(), point = guide_none()
       )
    )
  ) | (
    plot_coverage_quantiles(scores) + (
      plot_rel_score_by_quantile(relative_interval_score) +
      plot_bias(summary_performance)
    ) +
    plot_layout(heights = c(2, 1), guides = "collect")
  )
) +
  plot_layout(widths = c(1, 2), guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre", legend.box = "vertical"
  )
suppressWarnings(print(p))
```

@@

Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the Forecast Hub ensemble on the log scale though with greater uncertainty.
Visual differences were more marked on the natural scale with the surrogate model forecasting spuriously high peak incidence.
In a subset of example locations, we observed some variation in performance across locations, that the ensemble better-captured peak incidence, and that the surrogate model appeared biased toward overprediction.
Evaluating the relative performance of the surrogate model compared to the European Forecast Hub ensemble we found that the mean performance was substantially worse and that relative performance decreased with the forecast horizon.
The median forecast performance of the surrogate model was also worse when compared to forecasts from the ensemble though the majority of surrogate forecasts were within 50% of the performance observed for the ensemble forecast.
The difference in mean and median relative performance suggested a skewed distribution in scores, which we confirmed visually. 
This means that a relatively small fraction of forecasts were responsible for a substantial portion of the difference in performance.
Evaluating point forecast performance indicated a similar pattern of performance as that observed using the full predictive distribution though the relative performance of the surrogate model generally improved.
Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to incidence rates peaking across forecast locations linked to the spread of BA.2.
In general, the relative performance of the surrogate model degraded as forecast horizons increased with the distribution of relative performance having an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
The Forecast Hub ensemble was poorly calibrated, particularly at longer forecast horizons and larger prediction intervals, compared to the surrogate model though the surrogate model tended to be overly uncertain at large intervals.
The ensemble forecast was biased towards under prediction at short to medium forecast horizons but unbiased at longer horizons. 
In comparison, the surrogate model was biased towards over prediction and this bias increased linearly with the forecast horizon.

>>>>

---
title: Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble
preprint: true
author: 
  - name: Sam Abbott
    affiliation: 1
    corresponding: true
    email: sam.abbott@lshtm.ac.uk
  - name: Katharine Sherratt
    affiliation: 1
  - name: Nikos Bosse
    affiliation: 1
  - name: Hugo Gruson
    affiliation: 1
  - name:  Johannes Bracher
    affiliation: 2
  - name: Sebastian Funk
    affiliation: 1
affiliation:
  - code: 1
    address: London School of Hygiene and Tropical Medicine, London, UK
  - code: 2
    address: Chair of Statistics and Econometrics, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
abstract: > 
  Multi-model and multi-team ensemble forecasts have become widely used to generate reliable short-term predictions of infectious disease spread. Notably, various public health agencies have used them to leverage academic disease modelling during the COVID-19 pandemic. However, ensemble forecasts are difficult to interpret and require extensive effort from numerous participating groups as well as a coordination team. In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models. Here we used observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience to identify a set of properties present in current ensemble forecasts. We then developed a parsimonious forecast model intending to mirror these properties. We assess forecasts generated from this model in real time over six months (the 15th of January 2022 to the 19th of July 2022) and for multiple European countries. We focused on forecasts of cases one to four weeks ahead and compared them to those by the European forecast hub ensemble. We find that the surrogate model behaves qualitatively similarly to the ensemble in many instances, though with increased uncertainty and poorer performance around periods of peak incidence as measured by the Weighted Interval Score. The performance differences, however, seem to be partially driven by a subset of time points, and the proposed model appears better probabilistically calibrated than the ensemble. We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble, but more work is needed to understand the implicit epidemiological model that it represents. 
bibliography: library.bib
link-citations: true
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article
    extra_dependencies: ["float"]
    keep_tex: true
---

@@

Whilst there is robust and consistent evidence that multi-team ensemble forecasts provide reliable and performant forecasts across domains [@Reich2022-jo] they also have a range of downsides.
The most significant is the difficulty in interpreting them. This relates both to the underlying mechanisms for the forecasts they produce and to understanding if and when their behaviour is desirable. 
This impacts users' trust, how easily ensemble performance can be improved, and how easily contributor forecasts can be improved.
Forecasts from these ensembles also require considerable resource cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
Additional challenges with maintaining multi-team collaborations can include providing detailed feedback to those contributing forecasts that would allow them to improve their forecast approaches, providing incentives for forecasters to continue to contribute and adjust their models to changing conditions, and difficulty improving the quality of the ensemble by learning from past predictive performance [@Sherratt2022-ue].
Each of these issues may impact the long-term quality of the resulting forecasts and have implications for end-users. 
Little progress has so far been made in mitigating these downsides or in improving access to the high-quality and robust forecasts they seek to generate for geographies without coverage or for other infectious diseases. 
There has also been limited critical feedback on the structure of forecasting ensembling projects for infectious disease epidemiology and little evaluation of the effort required to produce them relative to their benefits for improving forecast performance.

In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysics [@Vernon2014-ov], emulation approaches have been used to circumvent resource requirement issues for complex models by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
These approaches generally take the same inputs as the models they seek to emulate and then are trained based on the output from those models.
In the context of epidemiological models, non-parametric emulation has been used to allow the rapid exploration of the parameter space of complex models that would otherwise be resource-prohibitive [@Iskauskas2022-fz; @Charles2022-ds].
Replicating forecasts from resource-intensive models may also be a promising avenue of future research. 
However, as an approach, it may be less useful for resolving some of the issues of multi-team and multi-model forecasts as it does not provide interpretability, key for stakeholder take-up. Additionally, it is not clear how the methods commonly used perform out of sample, and emulating a quantile-based forecast is an unaddressed methodological challenge as far as we are aware.

@@

To achieve these aims, we evaluate an initial attempt at developing a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles based on a set of clear assumptions.
We submitted this model to the European Forecast Hub and here we evaluate its performance relative to the Hub ensemble.
In this work, we first define the model and summarise its implementation, with a focus on minimal resource use and reproducibility as a GitHub Actions workflow [@GitHubActionsAbout].  
We then evaluate its real-time performance in comparison to the European Forecast Hub ensemble by visualising forecasts, using the weighted interval score [@bracherEvaluatingEpidemicForecasts2021], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where it performs less well.
Finally, we summarise our findings, discuss their implications, and highlight areas where more work is needed.
We aim for this work to highlight some of the potential implicit assumptions of current COVID-19 Forecast Hub ensembles, provide a sensible, low-resource, surrogate model where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.

@@

To understand the behaviour of the Forecast Hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team. 
Each hub targets a range of metrics, including test-positive reported cases, reported deaths, and hospitalisations; has a specific geographic focus, and asks for weekly forecasts (using epidemiological weeks i.e. Sunday to Saturday) over a time horizon of a few weeks.
Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
Here, we focus on reported cases and primarily on the European Forecast Hub but our observations hold, in our view, across COVID-19 Forecast Hubs and to a lesser degree targets.
We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator for other metrics such as hospitalisations [@Meakin2022-hw], and they are generally the most challenging to predict [@Sherratt2022-ue].
In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models, mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@Srivastava2020-qv; @Li2021-cj], semi-mechanistic approaches that blend both of these approaches [@Castro2021-lj; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-jn; @Adamik2020-if], and human insight based forecast models that may also include elements of other methods [@Karlen2020-pt; @Bosse2022-hm].
Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].

We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub [@Sherratt2022-ue; @EuroHub] from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth` and defined in the following section).
We included all locations covered by the European forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@JHU].
We used the latest available observed data as of the 1st of September 2022 (commit `f6922c3e4bdcb055abcbba8e73472afacac4cf40` from [@EuroHubGitHub]).
Incidence was aggregated by epidemiological week (i.e. Sunday through Saturday).
As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
To account for this we followed the practice of the European forecasting hub project in excluding forecasts made using anomalous truth data in the week of the forecasts production and excluding forecasts for target weeks with anomalous data [@Sherratt2022-ue].
We defined anomalous data based on the implementation used by [@EuroHub] where a data point is considered anomalous if a future revision alters it by more than 5%.

@@

We have contributed a range of forecasts to COVID-19 forecast collaborations, generally focused on semi-mechanistic statistical methods and human insight-based forecasts.
Our forecast submissions have not systematically over- or under-performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
Based on our observations our forecasts have generally captured the current trend relatively well but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns. 
Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European forecasting hub [@AzurePricing]. 
This resource usage is likely beyond the capacity of many interested in having access to state-of-the-art short-term forecasts of infectious diseases.
In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
We found that our human insight-based forecasts outperformed our model-based forecasts on average. 
This was particularly the case when forecasting cases and at longer forecast horizons.
We hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].

@@

The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models. Instead, we make use of a negative binomial observation model allowing the inclusion of overdispersion.
This choice is motivated by our belief that the underlying transmission process is an exponential discrete one and therefore a count error model with a log link function, where variance is linked to the mean, is a sensible choice. 
We suggest that part of the reason the hub ensembles exhibit such sharpness is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021]. Our set of assumptions and simplifications are summarised in Table \@ref(tab:assumptions).

@@

In our study period, incidence rates across European nations and in the UK were primarily driven by the spread of novel subvariants of concern related to the Omicron variant and changes in population susceptibility.
Many countries, such as the UK, saw large BA.1 waves in January, resulting in declining incidence rates through February (Figure \@ref(fig:vis-forecasts)).
From late February through to the end of May, most nations saw another wave driven by BA.2. This wave typically saw lower reported incidence rates, and was characterised by a lower peak than the BA.1 wave with a more gradual decrease in incidence.
The end of our study period was dominated by the gradual take-over of the BA.4/BA.5 subvariants that again had a lower peak and lower absolute growth rates. Unlike earlier periods in the pandemic, our study period did not see the use of new non-pharmaceutical interventions (NPIs) in response to increasing COVID-19 incidence in most locations.
In addition, ascertainment rates likely reduced over time in most locations due to reductions in routine testing and test availability. Whilst both the reduced use of NPIs and testing generally occurred across nations our study period also marked an increase in the heterogeneity of the response to the COVID-19 pandemic with nations changing policy at different times and to different degrees.
This is in contrast to the early COVID-19 pandemic response for which most nations took similar actions at similar times.

We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European forecasting hub (nations of the European Union, the European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
In total `r nrow(metadata)` forecasts were made across all locations, with `r forecast_dates` unique forecast dates and `r models` independent forecast models (including the European hub baseline model).
Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model.
Of the remaining models submitted `r length(local_models)` were submitted in only one location.
Single-location models were clustered in a few locations, particularly in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project [@Sherratt2022-ue]).
Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
`r nrow(partial_models)` models were submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.

@@

```{r vis-forecasts, fig.cap = "a.) Forecasts of notified test-positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom, by forecast horizon (one and four weeks) for the surrogate model (orange) and forecast ensemble (green). 30\\%, 60\\%, and 90\\% prediction intervals are shown. The black line and points are the notified cases as of the date of data extraction rather than those available at the time. b.) A replicate of a.) but with incidence rates on the log scale. c.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom on the log scale.", out.width = "98%", fig.height = 8, fig.width = 12}
p <- (
  plot_forecast_custom(forecasts, log = FALSE) |
  plot_forecast_custom(forecasts, log = TRUE) +
  labs(y = "") |
  plot_wis(wis, locs)
) +
  plot_layout(widths = c(2, 2, 1), guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre"
  )
suppressWarnings(print(p))
```

@@

```{r plot-coverage, fig.height = 8, fig.width = 8, out.width = "98%", fig.cap = "a.) Empirical coverage of the surrogate (orange) and ensemble (green) forecasts at the 90\\%, 60\\%, and 30\\% prediction intervals stratified by forecast horizon. Ideally, a well-calibrated forecast should have empirical coverage for a given prediction interval that equals the nominal level of the interval (i.e., 30\\%, 60\\% and 90\\%, respectively). b.) Empirical coverage by quantile for both the surrogate and ensemble forecasts. A well-calibrated forecast should have empirical quantiles that match the theoretical ones. The green area of this figure corresponds to conservative forecasts. c.) Median relative weighted interval score by quantile and forecast horizon for the surrogate forecast compared to the ensemble forecast. d.) The bias of the ensemble and surrogate forecasts stratified by horizon."}
p <- (
  (
    (
      plot_coverage_range(coverage, ranges) +
       guides(
        fill = guide_none(), point_color = guide_none(), point = guide_none()
       )
    )
  ) | (
    plot_coverage_quantiles(scores) + (
      plot_rel_score_by_quantile(relative_interval_score) +
      plot_bias(summary_performance)
    ) +
    plot_layout(heights = c(2, 1), guides = "collect")
  )
) +
  plot_layout(widths = c(1, 2), guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre", legend.box = "vertical"
  )
suppressWarnings(print(p))
```

@@

Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the Forecast Hub ensemble on the log scale though with greater uncertainty.
Visual differences were more marked on the natural scale with the surrogate model forecasting spuriously high peak incidence.
In a subset of example locations, we observed some variation in performance across locations, that the ensemble better-captured peak incidence, and that the surrogate model appeared biased toward overprediction.
Evaluating the relative performance of the surrogate model compared to the European Forecast Hub ensemble we found that the mean performance was substantially worse and that relative performance decreased with the forecast horizon.
The median forecast performance of the surrogate model was also worse when compared to forecasts from the ensemble though the majority of surrogate forecasts were within 50% of the performance observed for the ensemble forecast.
The difference in mean and median relative performance suggested a skewed distribution in scores, which we confirmed visually. 
This means that a relatively small fraction of forecasts were responsible for a substantial portion of the difference in performance.
Evaluating point forecast performance indicated a similar pattern of performance as that observed using the full predictive distribution though the relative performance of the surrogate model generally improved.
Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to incidence rates peaking across forecast locations linked to the spread of BA.2.
In general, the relative performance of the surrogate model degraded as forecast horizons increased with the distribution of relative performance having an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
The Forecast Hub ensemble was poorly calibrated, particularly at longer forecast horizons and larger prediction intervals, compared to the surrogate model though the surrogate model tended to be overly uncertain at large intervals.
The ensemble forecast was biased towards under-prediction at short to medium forecast horizons but unbiased at longer horizons. 
In comparison, the surrogate model was biased towards over prediction and this bias increased linearly with the forecast horizon.
